Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-07-24 manuscript/response_to_reviews.Rmd

---
title: "Response to reviews"
output: pdf_document

bibliography: ["book.bib",  "packages.bib"]
link-citations: yes
biblio-style: "apalike"
csl: new-phytologist.csl
---

@@

69 this is a very important point—microsite is crucial both for the ultimate height an individual grows to as well as its vulnerability. The generalization that on a global level large trees seem disproportionately vulnerable is often greeted with the skeptical observation that in specific cases a small tree died next to a large one that survived. However, the prediction is that *all else being equal* large individuals should be more vulnerable than smaller *conspecifics*. There is so much variation in microsite and other factors that the greater vulnerability of larger individuals should emerge clearly only in carefully controlled experiments where all else is indeed (virtually) equal, or with high sample sizes where variation in microsite etc should random out and the signal of generally higher vulnerability should come through.

**Thank you.**

73 The “greater root reach” point is almost universally appealed to—cites for these counterarguments would be useful.

**We now address this in the discussion: "Belowground, taller trees would tend to have larger root systems [@enquist_global_2002], but this does not necessarily imply that they have greater access to or reliance on deep soil-water resources that may be critical during drought. While tree size can correlate with the depth of water extraction [@brum_hydrological_2019], the linkage is not consistent. Shorter trees can vary broadly in the depth of water uptake [@stahl_depth_2013], and larger trees may allocate more to abundant shallow roots that are beneficial for taking up water from rainstorms [@meinzer_partitioning_1999]. Moreover, reliance on deep soil-water resources can actually prove a liability during severe and prolonged drought, as these can experience more intense water scarcity relative to non-drought conditions [@chitratarak_roots_2018]."**

### 2020-07-24 manuscript/climate_sensitivity_draft.Rmd
Fixed references and slightly edited final paragraph #117; #120

As climate change drives increasing drought in many of the world's forests [@trenberth_global_2014; @intergovernmental_panel_on_climate_change_climate_2015], the fate of forests and their climate feedbacks will be shaped by the biophysical and physiological drivers observed here. 
Our results show that taller, more exposed trees and species with less drought tolerant leaf traits will be most affected, at least in terms of growth during the drought year. 
Resilience and survival are both linked to resistance (**REFS**), implying that the same factors may influence these. 
Indeed, while the influence of $PLA_{dry}$ and $\pi_{tlp}$ on drought resilience and survival remains to be tested, taller trees have lower resilience [@trugman_tree_2018; @gillerot_tree_2020] and survival [@bennett_larger_2015; @stovall_tree_2019]. 
As climate change-driven droughts affect forests worldwide, there is likely to be a shift from mature forests with tall, buffering trees to forests with a shorter overall stature [@mcdowell_pervasive_2020]. 
At this point, species whose drought resistance relies in part on existence within a buffered microenvironment (e.g., *F. grandifolia*) could in turn become more susceptible. 
Here, the relative importance of tree height *per se* versus crown exposure becomes critical, shaping whether the dominant trees of shorter canopies are significantly more drought resistant because of their shorter stature, or whether high exposure makes them as vulnerable as the taller trees of the former canopy.
Studies disentangling the influence of height and exposure on drought tolerance will be critical to answering this question. 
Ultimately, distributions of tree heights and drought tolerance traits across broad moisture gradients suggest that forests exposed to more drought will ultimately shift towards shorter stature (**REFS**) and be dominated by species with more drought-tolerant traits (**REFS**).
Our study helps to elucidate the mechanisms behind these patterns, opening the door for more accurate forecasting of forest responses to future drought.

>>>>

As climate change drives increasing drought in many of the world's forests [@trenberth_global_2014; @intergovernmental_panel_on_climate_change_climate_2015], the fate of forests and their climate feedbacks will be shaped by the biophysical and physiological drivers observed here. 
Our results show that taller, more exposed trees and species with less drought-tolerant leaf traits will be most affected, at least in terms of growth during the drought year. 
Resilience and survival are both linked to resistance (**REFS**), implying that the same factors may influence these. 
Indeed, while the influence of $PLA_{dry}$ and $\pi_{tlp}$ on drought resilience and survival remains to be tested, taller trees have lower resilience [@trugman_tree_2018; @gillerot_tree_2020] and survival [@bennett_larger_2015; @stovall_tree_2019]. 
As climate change-driven droughts affect forests worldwide, there is likely to be a shift from mature forests with tall, buffering trees to forests with a shorter overall stature [@mcdowell_pervasive_2020]. 
At this point, species whose drought resistance relies in part on existence within a buffered microenvironment (e.g., *F. grandifolia*) could in turn become more susceptible. 
Here, the relative importance of tree height *per se* versus crown exposure becomes crucial, shaping whether the dominant trees of shorter canopies are significantly more drought resistant because of their shorter stature, or whether high exposure makes them as vulnerable as the taller trees of the former canopy.
Studies disentangling the influence of height and exposure on drought tolerance will be critical to answering this question. 
Ultimately, distributions of tree heights and drought tolerance traits across broad moisture gradients suggest that forests exposed to more drought will ultimately shift towards shorter stature (**REFS**) and be dominated by species with more drought-tolerant traits (**REFS**).
Our study helps to elucidate the mechanisms behind these patterns, opening the door for more accurate forecasting of forest responses to future drought.
